Finiteness, Cardinality, and Combinatorics in Homotopy Type Theory

### 2020-01-04 paper.tex
bib  etc

\input{preamble}
\title{Semirings and Semimodules: A Substrate for Computation}
\author{Donnacha Oisín Kidney}
\begin{document}
\maketitle
\listoftodos
\tableofcontents
\include{sections/maths}
\include{sections/type-theory}
\include{sections/implementation}
\include{sections/free-semimodule}
\include{sections/univalence-for-privacy}
\include{sections/combinatorics}
\include{sections/ordered-monoids-and-search}
\include{sections/semirings-on-the-gpu}
\end{document}

>>>>

\input{preamble}
\title{Finiteness, Cardinality, and Combinatorics in Homotopy Type Theory}
\author{Donnacha Oisín Kidney}
\begin{document}
\maketitle
\listoftodos
\tableofcontents
\newcommand{\pifun}[2]{\prod_{\left( \mathit{ #1 } : #2 \right)}}
\begin{quote}
  \textbf{Abstract}
  We explore five notions of finiteness in Homotopy Type Theory \cite{hottbook}.
  We prove closure properties about all of these notions, culminating in a proof
  that decidable Kuratowski-finite sets form a topos.

  We extend the definitions to include infinite types, developing a similar
  classification of countable types.

  We use the definition of finiteness to formalise \emph{species}, in much
  the same way as in \cite{yorgeyCombinatorialSpeciesLabelled2014}.
  A clear duality with containers
  \cite{abbottContainersConstructingStrictly2005} falls out naturally from our
  definition.

  We formalise our work in Cubical Agda
  \cite{vezzosiCubicalAgdaDependently2019}, and we implement a library for proof
  search (including combinators for level-polymorphic fully generic currying),
  and demonstrate how it can be used to both prove properties and synthesise
  full functions given desired properties.
\end{quote}
\section{Introduction}
Proofs in a constructive setting are more substantial things than their
classical counterparts.
Rather than just evidence for some fact, they are objects in their own right,
which we can manipulate and extract information from.

Sometimes, the extra detail can reveal interesting subtleties: fissures which
divide homogeneous-seeming definitions into a variety of related concepts.
Finiteness is a prime example of this phenomenon: in classical set theory, a
finite set is a finite set, and there's not much else that we can say.
In constructive type theory, we start the bidding at no fewer than \emph{four}
different predicates representing finiteness
\cite{spiwackConstructivelyFinite2010}, upped to eight in
\cite{firsovVariationsNoetherianness2016a}: all of which differ in significant
and interesting ways.

Other times, the extra baggage is unwanted.
Maybe we want to enforce that a particular piece of information should be
hidden, or treated as irrelevant: ``yes, I have proven that there is a
particular number with this property, but I don't want to reveal \emph{which}
number.''

This is where Homotopy Type Theory \cite{hottbook} can step in.
Through primarily a generalisation of equality, we gain access to a number of
exotic types which can, among other things, allow us to perform this ``hiding''.
In the context of finiteness, the act of hiding is worth studying in its own
right, and further delineates several new finiteness predicates which collapse
even in standard Martin-Löf type theory
\cite{martin-lofIntuitionisticTypeTheory1980}.

The other headline feature of HoTT's generalised equality is univalence: put
simply, it allows us to treat isomorphic types as equal, giving us in the
constructive world access to a technique that is commonplace classically.
It turns out that finite types have a lot to say about isomorphisms, and we will
rely heavily on this aspect of HoTT to work with them.
\subsection{Contributions}
In this work, we will explore finite types in Cubical Type Theory
\cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
types, species, and demonstrate their practical uses for proofs in dependently
typed programming languages.

In section~\ref{strong-finiteness}, we will explore the ``strong'' notions of
finiteness (i.e. those at least as strong as Kuratowski finiteness
\cite{kuratowskiNotionEnsembleFini1920}), with a special focus on cardinal
finiteness.
We will delineate some subtly different finiteness predicates, and prove the
relationship between each.
We will prove, then, that every discrete Kuratowski finite type is cardinal
finite, and subsequently will show that discrete Kuratowski finite sets form a
topos.

In section~\ref{species-and-containers}, we will redefine our finiteness
predicates using \emph{containers}, and prove that the new definitions are
equivalent to the old.
Using this, we will prove much stronger relationships between our finiteness
predicates and relation-based definitions of finiteness.
Finally, We will show how the species-container duality falls out naturally from
these new definitions.

In section~\ref{infinite-cardinalities}, we will extend our study of finite
types to infinite but countable types.
We will see that the finiteness predicates are mirrored with countable
counterparts, and we will prove closure under the Kleene star and plus.

Proofs of finiteness have well-known practical applications in
constructive mathematics \cite{firsovDependentlyTypedProgramming2015}.
In section~\ref{practical}, we build a library which exploits these uses in
Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, allowing automation of
complex proofs over finite types.
We frame this in terms of the principle of omniscience for finite types.
Thanks to the flexibility afforded to us by Cubical Type Theory, we are able to
go further than the usual examples of this kind of proof automation: as well as
proving properties about functions, we can synthesise functions whole-cloth from
their desired properties.
Through the unified interface for finite and countable types, we can
reuse the automation machinery for \emph{partial} proof search over infinite
search spaces.
Along the way, we extend the work in \cite{allaisGenericLevelPolymorphic2019} to
prove isomorphisms between the curried and uncurried forms of \(n\)-ary
dependent functions.
\section{Strong Finiteness} \label{strong-finiteness}
\begin{figure}[h]
  \centering
  \begin{tikzcd}[cramped, row sep=small]
    &
    \ar[ddddd, dash, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
    \text{Ordered} &
    \ar[ddddd, dash, dashed, start anchor={[yshift=4ex]}, end anchor={[yshift=-4ex]}] &
    \text{Unordered}
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \ar[rrrr, dash, start anchor={[xshift=-6ex]}, end anchor={[xshift=10ex]}] &
    &
    &
    &
    {}
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    &
    &
    \text{Split Enumerable}
      \; \ref{split-enumerable}
      \ar[d, xshift=-1ex, bend right=30]
      &
    &
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \text{Discrete} &
    &
    \text{Manifest Bishop}
      \ar[u, xshift=1ex, bend right=30]
      \ar[rr, end anchor=south west, start anchor=south east, bend right=10, crossing over]
      \; \ref{manifest-bishop}
      &
    &
    \text{Cardinal}
      \; \ref{cardinal}
      \ar[ll, start anchor=north west, end anchor=north east, bend right=10, "\text{linOrd}" description]
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \ar[rrrr, dash, dashed, start anchor={[xshift=-6ex]}, end anchor={[xshift=10ex]}] &
    &
    &
    &
    {}
    \\ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \text{Non Discrete} &
    {} &
    \text{Manifest Enumerable}
      \; \ref{manifest-enumerability}
      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
      \ar[uu, bend right=30,  "\text{Discrete}" description]
      \ar[rr, start anchor=south east, end anchor=south west, bend right=10, crossing over]
      &
    {}
    &
    \text{Kuratowski}
      \; \ref{kuratowski}
      \ar[ll, start anchor=north west, end anchor=north east, bend right=10,  "\text{linOrd}" description]
      \ar[uu, xshift=1ex, bend right=30,  "\text{Discrete}" description]
      \ar[from=uu, xshift=-1ex, bend right=30, crossing over]
  \end{tikzcd}
  \caption{Classification of the strong definitions of finiteness, according to
    whether they are discrete (imply decidable equality) and whether they induce
    a linear order.}
  \label{finite-classification}
\end{figure}

In this section, we will explore five predicates for finiteness, classified in
figure~\ref{finite-classification}.
Of these, manifest enumerability is new, to our knowledge.

Figure~\ref{finite-classification} organises the predicates according to
their ``strength''; i.e. how much information they provide about a conforming
type.
For instance, a proof that some type \(A\) is manifestly Bishop finite (the
strongest of the notions, explored in section~\ref{manifest-bishop}) also tells
us that \(A\) is discrete (has decidable equality), and gives us a linear order
on the type.
A type that is Kuratowski finite (section~\ref{kuratowski}) has no such extra
features: indeed, we will see examples of Kuratowski finite types which are not
even sets, never mind discrete ones.

In the following sections, we will go through each of the predicates, proving
how to weaken each (i.e. we will provide a proof that every cardinally finite
type is Kuratowski finite), and how to strengthen them, given the required
property.
In terms of figure~\ref{finite-classification}, this amounts to providing proofs
for each arrow.

For each predicate, we will also prove its closure over sums and products in
both dependent and non-dependent forms, if such a closure exists.
This will culminate in our main result for this section: the formal proof that
decidable Kuratowski finite sets form a topos. \todo[noline]{reference for this}
More specifically, we show that cardinal finite sets form a topos, that
decidable Kuratowski finite sets are equivalent in strength to cardinal finite
sets, and carry the proof over.
This result relies on proofs on each of the other finiteness predicates.
\subsection{Split Enumerability} \label{split-enumerable}
We will start with the simplest definition of finiteness: we say a set is
enumerable if there is a list of its elements which contains every element in
the set.
More formally:
\begin{definition}[Split Enumerable Set]
  \begin{equation}
    \mathcal{E}!(A) = \sum_{(\mathit{xs} : \textbf{List} A)} , \prod_{(x : A)} , x \in xs
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
\end{definition}

The term \(x \in \mathit{xs}\) hides a subtle detail of this definition.
Depending on how we choose to represent \(x \in \mathit{xs}\), the resulting
predicate will be equivalent in strength to either manifest Bishop finiteness
(section~\ref{manifest-bishop}) or manifest enumerability
(section~\ref{manifest-enumerability}).
For now, we will define it like so:

\begin{definition}[List Membership] \label{list-membership}
  \begin{equation}
    x \in \mathit{xs} = \sum_{(i : \mathbf{Fin}(\text{length}(\mathit{xs})))} , \mathit{xs} \ind i \equiv x
  \end{equation}
  Here, \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\),
  and \(\ind\) is an indexing function.
  In other words, to say that some item \(x\) is in some list \(\mathit{xs}\),
  we say that there is an index into \(\mathit{xs}\) which selects an element of
  \(\mathit{xs}\) which is equivalent to \(x\).
\end{definition}

Clearly, this definition of list membership tells us not just that an item is in
a list, but also its \emph{position} in that list.
That piece of information gives use decidable equality, restricting the
predicate to discrete sets.
In section~\ref{manifest-enumerability}, we will show how to hide the position,
giving us a variant of the predicate which allows for non-set finite types.
\subsubsection{Decidable Equality}
We will now prove that split enumerability implies decidable equality.
We first need to prove a lemma regarding surjections to split enumerable types.
\begin{lemma} \label{split-enum-surj}
  Any split enumerable type has a split surjection from a finite prefix of the
  natural numbers.
\end{lemma}
\begin{proof} 
  Let \(E\) be a proof of type \(\mathcal{E}!(A)\) for some type \(A\), and
  \(n\) be the length of the support list of \(E\).
  Then let \(f : \textbf{Fin}(n) \rightarrow A\) be defined as the indexing
  function into the support list of \(E\).

  We then prove this function to be a split surjection by providing a fiber over
  all of the elements of its domain: this fiber is the second component of \(E\).

  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{to-surj}
\end{proof}

Because of careful arrangement of previous definitions, this proof is only three
lines of Agda.
It is not difficult to derive decidable equality from this fact.
\begin{lemma} \label{split-enum-discrete}
  Any split enumerable type has decidable equality (is discrete).
\end{lemma}
\begin{proof}
  We use a corollary, proven in the appendix, that if there is a
  split-surjection from \(A\) to \(B\), and \(A\) is discrete, then \(B\) is
  also discrete.
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{discrete}
\end{proof}

This also tells us that any split enumerable type must be a set.
\begin{lemma}
  Any split enumerable type is a set.
\end{lemma}
\begin{proof}
  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, since split
  enumerable types have decidable equality
  (proposition~\ref{split-enum-discrete}), they are sets.
\end{proof}

Though it's not strictly needed for the rest of the construction, it is
instructive to also prove inverse regarding surjections; i.e. that we can
derive \(\mathcal{E}!\) from a surjection from another finite type.
\begin{lemma} \label{split-enum-from-surj}
  Any type with a split surjection from another split enumerable type is itself
  split enumerable.
\end{lemma}
\begin{proof}
  Let \(E\) be a proof of split enumerability for some set \(A\), and let \(f\)
  be a surjection from \(A\) to \(B\).
  The support list for the proof of split enumerability over \(B\) can be
  obtained by mapping the surjective function over the support list of \(E\).

  To obtain the cover proof, let \(f^{-1}\) be the inverse function of \(f\)
  (which we can obtain because it is a split-surjection).
  Using the cover proof of \(E\), we can show that, for all \(y : B\), \(f^{-1}(y)\)
  is in the support list for \(E\).
  Using the following lemma (proven in the appendix):
  \begin{equation}
    x \in xs \rightarrow f (x) \in \text{map} \; f \; \mathit{xs}
  \end{equation}
  We obtain a proof that:
  \begin{equation}
    \forall (y : B). f (f^{-1}(y)) \in \text{map} \; f \; \mathit{xs}
  \end{equation}
  And, since \(f (f^{-1}(y)) \equiv y\) (by the definition of a surjection), we
  have proven our goal.

  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{from-surj}
\end{proof}
\subsubsection{Closure}
In this section we will prove closure under various operations for split
enumerable sets.
We are working towards a topos proof, which requires us to prove closure under
a variety of operations: for now, we only have enough machinery to demonstrate
the semiring operations, and dependent sums.
in order to show closure under exponentials (function arrows), we will need an
equivalence with \(\mathbf{Fin}\), which will be provided in
section~\ref{manifest-bishop}.
\begin{lemma}
  \(\bot\) is split enumerable.
\end{lemma}
\begin{proof}
  The support list for \(\bot\) is the empty list.
  The cover proof is present via the principle of explosion.
\end{proof}
\begin{lemma}
  \(\top\) is split enumerable.
\end{lemma}
\begin{proof}
  The support list for \(\top\) is the singleton list containing \(\text{tt}\).
  The cover proof is a constant function which returns an index into the first
  item in the list.
\end{proof}
Both of those cases were relatively simple.
Next, we will look into how to combine proofs of split enumerability.
\begin{theorem}
  Split-enumerability is closed under \(\sum\).
\end{theorem}
\begin{proof}
  Let \(E_A\) be a proof of split enumerability for some type \(A\), and \(E_U\)
  be a function of the type:
  \begin{equation}
    E_U : \prod_{(x : A)} , \mathcal{E}!(U(x))
  \end{equation}
  In other words, a function which returns a proof of split enumerability for
  each member of the family \(U\).

  To obtain the support list, we concatenate the support lists of all the proofs
  of split-finiteness for \(U\) over the support list of \(E_A\).
  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-enum}
  ``do-notation'' is available to us as we're working in the list monad.

  The cover proof follows in a simple, if tedious, way.
\end{proof}
\begin{lemma}
  Split-enumerability is closed under Cartesian product (non-dependent product).
\end{lemma} 
\begin{proof}
  Since non-dependent products are simply a special case of the \(\sum\) type,
  we can reuse the proof of closure over \(\sum\) here.
\end{proof}
\begin{lemma}
  Split-enumerability is closed under disjoint union (non-dependent sum).
\end{lemma}
\begin{proof}
  Again, this can be derived from closure of the \(\sum\) type.
  Disjoint union between two types \(A\) and \(B\) can be represented by the
  following type:
  \begin{equation}
    A + B = \sum_{(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
  \end{equation}
  Then, since all of \(\mathbf{Bool}\), \(A\), and \(B\) are split enumerable,
  the type \(A + B\) is split enumerable.
\end{proof}
The final case can be derived from the previous closures; we give it here anyway
for clarity.
\begin{lemma} \label{fin-split-enum}
  Any finite prefix of the natural numbers (what we have been calling
  \(\textbf{Fin}\)) is split enumerable.
\end{lemma}
\begin{proof}
  For the support list, we will make use of the function tabulate, which returns
  a list consisting of the application of \(f\) to every \(\textbf{Fin}(n)\),
  given \(n\).
  \begin{equation}
    \text{tabulate}(n , f) = \left[ f(0) , f(1) \dots f(n - 1) \right]
  \end{equation}
  In Agda:
  \ExecuteMetaData[agda/Data/List.tex]{tabulate}
  Tabulation with the identity function gives us our cover list.

  The cover proof is obviously true given the definition of tabulate above; in
  Agda it requires a little more work to show.
  \ExecuteMetaData[agda/Data/List/Membership.tex]{fin-in-tabulate}
\end{proof}
\subsection{Manifest Bishop Finiteness} \label{manifest-bishop}
\begin{definition}[Manifest Bishop Finiteness]
  \begin{equation}
    \mathcal{B}(A) = \sum_{(\mathit{xs} : \textbf{List}(A))} , \prod_{(x : A)} , x \inunique xs
  \end{equation}
\end{definition}
The only difference between this predicate and split enumerability is the list
membership term: we use \(\inunique\) here, where \(x \inunique \mathit{xs}\) is
to be read as ``\(x\) occurs exactly once in \(\mathit{xs}\)''.
It is defined in terms of our previous list membership term
(definition~\ref{list-membership}).
\begin{definition}[Unique List Membership]
  We say an item \(x\) is ``uniquely in'' some list \(\mathit{xs}\) if its
  membership in that list is a \emph{contraction}; i.e. its membership proof
  exists, and all such proofs are equal.
  \begin{equation}
    x \inunique \mathit{xs} = \text{isContr}(x \in \mathit{xs})
  \end{equation}
\end{definition}

A nice consequence of prohibiting duplicates is that now the length of the
support list is the same as the cardinality of the set.
\subsubsection{Equivalence}
While manifest Bishop finiteness may seem stricter than split enumerability, we
will see soon that it is actually equivalent in strength.
Why study it separately to split enumerability at all, then?
What purpose does it serve on our road to Kuratowski finiteness?
Those questions are answered by the following theorem:
\begin{lemma} \label{bishop-equiv}
  Any manifestly Bishop finite type has an equivalence with a finite prefix of
  the natural numbers.
\end{lemma}
\begin{proof}
  There are many equivalent definitions of equivalence in HoTT.
  Here we take the version preferred in the Cubical Agda library: contractible
  maps \cite[Definition~4.4.1]{hottbook}. \todo[noline]{Need this in appendix also}
  Because of the parallels between contractible maps and split surjections,
  the proof proceeds much the same as \ref{split-enum-surj}.

  First, we define the map as the indexing function into the support list.
  Proving this map is a contraction is precisely the cover proof.

  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{equiv}
\end{proof}

As we can see, our list-based definition of manifest Bishop finiteness is really
an equivalence with \(\mathbf{Fin}\) in disguise.
This will be the key component in establishing closure under exponentials.

We would like to draw attention to the ``power-to-weight ratio'' of this proof.
We have mentioned how important the equivalence will be for our later proofs: in
general, the ability to transport via univalence opens up many opportunities

In Cubical Type Theory, equivalences are much more useful than their counterpart
(isomorphisms) in Martin-Löf  Type Theory: arbitrary proofs and functions can be
\emph{transported} along equivalences, giving a constructive justification for
an intuitive and useful technique which is commonplace classically.
The proof itself, however, is three lines long: thanks to careful definitions,
the fact that manifest Bishop finiteness is the ``same'' as equivalences to
\(\mathbf{Fin}\) becomes obviously true.
Later we will see, by defining lists as containers, a proof that manifest Bishop
finiteness is equivalent (in the formal sense) to equivalences with
\(\mathbf{Fin}\). \todo[noline]{This needs to be added to the containers
  section.}
\todo[noline]{More on this? }
\subsubsection{Relationship to Split Enumerability}
We now show that manifest Bishop finiteness has equal strength to split
enumerability.
\begin{lemma}
  Any manifest Bishop finite set is split enumerable.
\end{lemma}
\begin{proof}
  The support set carries over simply, and the cover proof can be taken from the
  first component of the cover proof from the proof of manifest Bishop
  finiteness.
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{bishop-to-split-enum}
\end{proof}
\begin{theorem} \label{split-enum-to-manifest-bishop}
  Any split enumerable set is manifest Bishop finite.
\end{theorem}
\begin{proof}
  Let \(E\) be a proof of split enumerability for some set \(A\).
  From proposition~\ref{split-enum-discrete} we can derive decidable equality on
  \(A\), and using this we can define a function \(\text{uniques}\) which filters
  out duplicates from lists of \(A\)s.
  \begin{equation}
    \text{uniques} : \textbf{List}(A) \rightarrow \textbf{List}(A)
  \end{equation}
  This gives us our support list.

  It suffices now to prove the following:
  \begin{equation}
    \prod_{(x : A)} , \prod_{(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
  \end{equation}
  And from that we can generate our cover proof.

  Full details of the formal proof are provided in the appendix.
\end{proof}
Note that while manifest Bishop finiteness as split enumerability are equivalent
in ``strength'', the two types are not equivalent.
In particular, there are infinitely many inhabitants of \(\mathcal{E}!\), while
for a type \(A\) with \(n\) inhabitants, there are only \(n!\) inhabitants of
\(\mathcal{B}(A)\).
\subsubsection{Closure}
Proving equal strength of split enumerability and manifest Bishop finiteness
allows us to carry all of the previous proofs of closure over to manifest Bishop
finite sets (and vice-versa).
Missing from our previous proofs was a proof of closure of functions.
We remedy that here.
\begin{theorem}
  Manifest bishop finiteness is closed over dependent functions (\(\prod\)-types).

  Formally, given a type \(A\) and a type family \(B\) on \(A\), when \(A\) is
  manifestly Bishop finite:
  \begin{equation}
    \mathcal{B}\left( A \right)
  \end{equation}
  And \(B\) is manifestly Bishop finite over all points of \(A\):
  \begin{equation}
    \prod_{(x : A)} , \mathcal{B}\left( B(x) \right)
  \end{equation}
  Then we have the following:
  \begin{equation}
    \mathcal{B}\left(\prod_{(x : A)} , B(x)\right)
  \end{equation}
\end{theorem}
\begin{proof}
  This proof is essentially the composition of two transport operations, made
  available to us via univalence.

  First, we will simplify things slightly by working only with split
  enumerability.
  As this is equal in strength to manifest Bishop finiteness, any closure proofs
  carry over.

  Secondly, we will replace \(A\) in all places with \(\mathbf{Fin}(n)\).
  Since we have already seen an equivalence between these two types, we are
  permitted to transport along these lines.
  This is the first transport operation.

  The bulk of the proof now is concerned with proving the following:
  \begin{equation}
    \left( \prod_{(x : \mathbf{Fin}(n))} , \mathcal{E}!(A(x)) \right) \rightarrow \mathcal{E}!\left( \prod_{(x : \mathbf{Fin}(n))} , A(x) \right)
  \end{equation}
  Our strategy to accomplish this will be to consider functions from
  \(\mathbf{Fin}(n)\) as \(n\)-tuples over some type family \(T\).
  \begin{equation}
    \begin{aligned}
      \mathbf{Tuple}(T, 0)   &= \top \\
      \mathbf{Tuple}(T, n+1) &= A(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
    \end{aligned}
  \end{equation}
  This type is manifestly Bishop finite, as it is constructed only from products
  and the unit type.

  We then prove an isomorphism between this representation and \(\Pi\)-types.
  \begin{equation}
    \mathbf{Tuple}(T, n) \iff \prod_{(x : \mathbf{Fin}(n))} , T(x)
  \end{equation}
  This allows us to transport our proof of finiteness on tuples to one on
  functions from \(\mathbf{Fin}\) (our second transport operation), proving our
  goal.

  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{pi-clos}
\end{proof}
\subsection{Manifest Enumerability} \label{manifest-enumerability}
As we said in section~\ref{split-enumerable}, we will now retrieve a distinction
between enumerability and manifest Bishop finiteness.
We do so by propositionally truncating the membership proof:
\begin{definition}[Manifest Enumerability]
  \begin{equation}
    \mathcal{E}(A) = \sum_{(\mathit{xs} : \mathbf{List}(A))} , \prod_{(x : A)} , \lVert x \in \mathit{xs} \rVert
  \end{equation}
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{def}
\end{definition}

There is another definition, which lies between manifest enumerability and
bishop finiteness, where we would truncate just the equality, but reveal the
position in the membership proof.
\begin{equation}
  \mathcal{E}'(A) = \sum_{(\mathit{xs} : \mathbf{List}(A))} , \prod_{(x : A)} , \sum_{(i : \textbf{Fin}(\text{length}(\mathit{xs})))} , \lVert \mathit{xs} \ind i \equiv x \rVert
\end{equation}
However, this simply amounts to a proof of manifest Bishop finiteness over
set-truncated types, rather than a significantly different construction, so it
does not interest us here.
\subsubsection{Higher Homotopy Levels}
By hiding the position, we have essentially removed the ``decidable'' component
from split enumerability.
Our predicate now becomes general enough to work with non-sets: we will show
here that the circle is manifestly enumerable.
\begin{theorem}
  The circle \(S^1\) is manifestly enumerable.
\end{theorem}
\begin{proof}
  As the cover proof is a truncated proposition, we need only consider the point
  constructors, making this poof the same as the proof of split enumerability on
  \(\top\).

  The support list will be a singleton list containing the single point in the
  type, and the cover proof will be an index pointing at the first element in
  the support list.

  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{circ-inst}
\end{proof}
\subsubsection{Surjections}
This predicates relation to surjectivity is much the same as split
enumerability's relation to \emph{split} surjectivity.
Deriving a surjection from \(\mathbf{Fin}\) is straightforward:
\begin{lemma}
  If a type \(A\) is manifestly enumerable, then there is a surjection from
  \(\mathbf{Fin}\) into \(A\).
\end{lemma}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}
And building from a surjection is similarly not too difficult:
\begin{lemma}
  If there is a surjection from a manifestly enumerable type \(A\) to another
  type \(B\), then \(B\) is manifestly enumerable.
\end{lemma}
\begin{proof}
  The proof proceeds in much the same way as in
  proposition~\ref{split-enum-from-surj}.
  The only difference is that the second component contains a truncation, but
  since we can always eliminate from a truncation into a truncation this is not
  a problem.

  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{from-surj}
\end{proof}
\subsubsection{Relation to Split Enumerability}
As split enumerability is stronger than manifest enumerability, we can convert
from one to the other easily.
\begin{lemma}
  Any split enumerable type is also manifestly enumerable.
\end{lemma}
\begin{proof}
  The proof carries over via truncation of the cover proof.
\end{proof}
The main missing piece from manifest enumerability is decidable equality: by
truncating the membership proof, we have removed the ability to distinguish
(decidably) between members of the type.
Indeed, resupplying decidable equality is precisely sufficient to recover split
enumerability.

\begin{theorem} \label{manifest-enum-to-split}
  A manifestly enumerable type with decidable equality is split enumerable.
\end{theorem}
\begin{proof}
  The support list stays the same between both enumerability proofs.

  For the cover proof, we need a way to get the contents out of a truncated
  value.
  We can do exactly that with the following lemma, called recompute:
  Given a decision procedure for some type \(A\), and a truncation for the
  proposition \(A\), we can discount the possibility of \(A\) being false, and
  therefore extract the true decision.

  For the cover proof, our obligation now becomes constructing a decision for
  membership of the support list.
  This is straightforward given decidable equality.

  \ExecuteMetaData[agda/Cardinality/Finite/ManifestEnumerable.tex]{to-split-enum}
\end{proof}
By theorem~\ref{split-enum-to-manifest-bishop}, we also can derive that any
manifestly enumerable type with decidable equality is Bishop finite.
\subsubsection{Closure}
\begin{lemma}
  Manifest enumerability is closed under dependent sum, disjoint union
  (non-dependent sum), and Cartesian product (non-dependent product).
\end{lemma}
\begin{proof}
  For these three closures, the proofs on split enumerability consisted of a
  list manipulation followed by a proof that membership was preserved by the
  list manipulation.
  Because we separate these two concerns, the proofs carry over onto manifest
  enumerability: the support list manipulation stays the same, and the
  cover proofs are performed ``under'' the truncation.
\end{proof}
Notice that we do not have closure under functions: without decidability,
manifest enumerability is not closed under function arrows.
\subsection{Cardinal Finiteness} \label{cardinal}
We now turn our attention to notions of finiteness which do not induce a linear
order.
\begin{definition}[Cardinal Finiteness]
  A type \(A\) is cardinally finite, \(\mathcal{C}\), if it has a
  propositionally-truncated proof of bishop finiteness.
  \begin{equation}
    \mathcal{C}(A) = \lVert \mathcal{B}(A) \rVert
  \end{equation}
\end{definition}
\subsubsection{Characterisation}
Though we have chosen Bishop finiteness as our underlying finiteness proof to
truncate, we could have in fact used any equal-strength proof
(lemma~\ref{equal-strength}).
Because this predicate is a mere proposition, what we called before ``equal
strength'' now implies full equivalence.
With this proven, we can use any one of our equal-strength finite definitions
(\(\mathcal{E}!\), \(\mathcal{B}\)).
In particular, we have:
\begin{lemma}
  Cardinal finiteness is equivalent to a propositionally truncated equivalence
  with some \(\textbf{Fin}\).
  \begin{equation}
    \mathcal{C}(A) \simeq \lVert \sum_{(n : \mathbb{N})} , \textbf{Fin}(n) \simeq A \rVert
  \end{equation}
\end{lemma}
\begin{proof}
  By proposition~\ref{equal-strength}, we only need to provide conversion
  functions between the two sides of the equivalence.
  For left to right, we have proposition~\ref{bishop-equiv}.
  For right to left, we can transport along the equivalence, meaning we need
  only prove that \(\textbf{Fin}(n)\) is Bishop finite.
  We know this to be true, since it is split enumerable
  (proposition-\ref{fin-split-enum}).
\end{proof}
\subsubsection{Closure}
The closure proofs for cardinal finiteness are especially easy.
In contrast to manifest enumerability, under the propositional truncation we
have a full proof of bishop finiteness, meaning that all of the closure proofs
carry over.
\begin{lemma}
  Cardinal finiteness is closed under dependent and non-dependent sums,
  products, and functions.
\end{lemma}
\begin{proof}
  All closure functions can be lifted under propositional truncation.
  Therefore, cardinal finiteness has the same closure properties as manifest
  bishop finiteness.
\end{proof}
\subsubsection{Strength}
We can eliminate from cardinal finiteness in a coherently constant way:
effectively, anything which doesn't ``notice'' the order of the internal support
list is usable.

First, even though the cardinality of the type is hidden under the truncation,
we can pull it out, as it is constant through permutations of the list.
\begin{theorem}
  Given a cardinally finite type, we can derive the type's cardinality, as well
  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
  the same cardinality.
  \begin{equation}
    \mathcal{C}(A) \rightarrow \sum_{(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
  \end{equation}
\end{theorem}
\begin{proof}
  We eliminate from the proof of cardinal finiteness via the eliminator
  described in proposition~\ref{elim-non-prop}.
  We pass a function \(\text{alg}\) which preserves the first part of the pair,
  and truncates the second.
  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{trunc-alg}

  It remains only to show that this function is coherently constant.
  \begin{equation}
    \prod_{(x : \sum_{(n : \mathbb{N})} , \textbf{Fin}(n) \simeq A)} ,
    \prod_{(y : \sum_{(m : \mathbb{N})} , \textbf{Fin}(m) \simeq A)} ,
    \text{alg}(x) \equiv \text{alg}(y)
  \end{equation}

  We will first address the first component of the output.
  To show that \(n\) and \(m\) are equal, we first observe that
  \begin{equation}
    \textbf{Fin}(n) \equiv \textbf{Fin}(m)
  \end{equation}
  since both are equivalent to \(A\).
  Relying on the fact that \(\textbf{Fin}\) is injective
  (proposition~\ref{fin-inj}), we can derive that \(n\) and \(m\) are equal.

  Since the second components of the output are propositions, they are
  definitionally equal.
  \ExecuteMetaData[agda/Cardinality/Finite/Cardinal.tex]{const-alg}
\end{proof}

\begin{lemma} \label{fin-inj}
  \(\textbf{Fin}\) is injective.
\end{lemma}
\begin{proof}
  Let \(n\) and \(m\) be natural numbers.
  We want to derive a proof of \(n \equiv m\) from \(\textbf{Fin}(n) \equiv
  \textbf{Fin}(m)\).

  We prove by contradiction (since the property we are interested in is
  decidable, this is valid constructively).
  When \(n\) does not equal \(m\), there are two possible cases.
  \begin{equation}
    \neg (n \equiv m) \implies
    \begin{cases}
      m \equiv 1 + n + k, n < m  \\
      n \equiv 1 + m + k, m < n
    \end{cases}
  \end{equation}
  Take the first case, without loss of generality.
  \todo[inline]{Finish prose for this proof}

  The Agda version of this proof is in the appendix (\ref{agda-fin}).
\end{proof}

Another property which does not rely on an internal order is decidable equality.
\begin{theorem} \label{cardinal-finite-discrete}
  Any cardinal-finite set has decidable equality.
\end{theorem}
\begin{proof}
  We first show that decidable equality is a proposition itself.
  We know that if a type \(A\) is a proposition, then the decision over that
  type is also a proposition.
  Then, via Hedberg's theorem, we know that any type with decidable equality is
  a set, meaning that paths in that type are themselves propositions.
  Therefore we can derive that a decision of equality on elements with decidable
  equality is a proposition, and by function extensionality we see that
  decidable equality is itself a proposition.
